---
title: "Group 5, R Project 3"
author: "Kera Whitley, Laura Mathews, Nataliya Peshekhodko"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'

abstract: "This is abstarct."

output: pdf_document
---
```{r message=FALSE}
library(dplyr)
library (ggplot2)
library (gridExtra)
```

# First part


## 1.1. Read data
Read the data from txt file.

```{r}
data <- read.delim("mpg.txt", header=TRUE, sep=" ")
```


## 1.2. Two sample t-test

T-test with equal variance 

```{r}
equal.var.test <- t.test( MPG~Country, data = data, var.equal=TRUE, conf.level=0.95)
equal.var.test
```

Based on the equal variance t.test **p-value=`r equal.var.test$p.value`**, we can reject the null hypothesis. 

T-test with unequal variance: 

```{r}
unequal.var.test <- t.test( MPG~Country, data = data, var.equal=FALSE, conf.level = 0.95)
unequal.var.test
```

Based on the unequal variance t.test **p-value=`r unequal.var.test$p.value`**, we also can reject the null hypothesis. 

## 1.3. How well the normality assumption met by the data

Checking data normality

```{r message=FALSE}
us <- filter(data, Country == "US")
japan <- filter(data, Country == "Japan")

par(mfrow=c(2,2))
hist_us<- ggplot (us, aes(MPG))+geom_histogram(aes(y=..density..),colour="black", fill = "white") + 
  stat_function(fun=dnorm, args = list (mean = mean(us$MPG), sd=sd(us$MPG)), colour = "red")
qq_us <- ggplot(us, aes(sample = MPG))+stat_qq()+stat_qq_line()
grid.arrange(hist_us, qq_us, ncol=1, nrow=2)

hist_jpn <- ggplot (japan, aes(MPG))+geom_histogram(aes(y=..density..),colour="black", fill = "white") + 
  stat_function(fun=dnorm, args = list (mean = mean(japan$MPG), sd=sd(japan$MPG)), colour = "red")
qq_jpn <- ggplot(japan, aes(sample = MPG))+stat_qq()+stat_qq_line()
grid.arrange(hist_jpn, qq_jpn, ncol=1, nrow=2)
```


## 1.4. Preferable test for selected dataset

# Second part



```{r}


get_power <- function (n1, n2, var1, var2) {
  
  alpha = 0.05
  N = 1000
  mu_1 = -5
  mu_2 = 0
  mu1_mu2 = -5
  A_sample <- as.data.frame ( replicate (N, rnorm (n1, mean = mu_1, sd = sqrt (var1))) )
  B_sample <- as.data.frame ( replicate (N, rnorm (n2, mean = mu_2, sd = sqrt (var2))) )
  
  pvalues_eq <- c()
  pvalues_noteq <- c()
  pwr_eq_sigma <- c()
  pwr_noteq_sigma <- c()
  
  for (col in 1:ncol(A_sample)) {
  pvalues_eq <-append (pvalues_eq, t.test (A_sample[,col], B_sample[,col], var.equal=TRUE, 
                                           conf.level = 0.95)$p.value )
  pvalues_noteq <-append( pvalues_noteq, t.test (A_sample[,col], B_sample[,col], var.equal=FALSE, 
                                                 conf.level = 0.95)$p.value )}
  
  pwr_eq_sigma <- append (pwr_eq_sigma, sum (pvalues_eq < alpha)/N)
  pwr_noteq_sigma <-append (pwr_noteq_sigma,  sum (pvalues_noteq < alpha)/N )

  
  mu_1 = -1
  mu_2 = 0
  mu1_mu2 = -1
  A_sample <- as.data.frame ( replicate (N, rnorm (n1, mean = mu_1, sd = sqrt (var1))) )
  B_sample <- as.data.frame ( replicate (N, rnorm (n2, mean = mu_2, sd = sqrt (var2))) )
  pvalues_eq <- c()
  pvalues_noteq <- c()
  for (col in 1:ncol(A_sample)) {
  pvalues_eq <-append (pvalues_eq, t.test (A_sample[,col], B_sample[,col], var.equal=TRUE, 
                                           conf.level = 0.95)$p.value )
  pvalues_noteq <-append( pvalues_noteq, t.test (A_sample[,col], B_sample[,col], var.equal=FALSE, 
                                                 conf.level = 0.95)$p.value )}
  pwr_eq_sigma <- append (pwr_eq_sigma, sum (pvalues_eq < alpha)/N)
  pwr_noteq_sigma <-append (pwr_noteq_sigma,  sum (pvalues_noteq < alpha)/N )

  #Added this section
  mu_1 = 0
  mu_2 = 0
  mu1_mu2 = 0
  A_sample <- as.data.frame ( replicate (N, rnorm (n1, mean = mu_1, sd = sqrt (var1))) )
  B_sample <- as.data.frame ( replicate (N, rnorm (n2, mean = mu_2, sd = sqrt (var2))) )
  pvalues_eq <- c()
  pvalues_noteq <- c()
  for (col in 1:ncol(A_sample)) {
    pvalues_eq <-append (pvalues_eq, t.test (A_sample[,col], B_sample[,col], var.equal=TRUE, 
                                             conf.level = 0.95)$p.value )
    pvalues_noteq <-append( pvalues_noteq, t.test (A_sample[,col], B_sample[,col], var.equal=FALSE, 
                                                   conf.level = 0.95)$p.value )}
  pwr_eq_sigma <- append (pwr_eq_sigma, sum (pvalues_eq < alpha)/N)
  pwr_noteq_sigma <-append (pwr_noteq_sigma,  sum (pvalues_noteq < alpha)/N )
  
  mu_1 = 1
  mu_2 = 0
  mu1_mu2 = 1
  A_sample <- as.data.frame ( replicate (N, rnorm (n1, mean = mu_1, sd = sqrt (var1))) )
  B_sample <- as.data.frame ( replicate (N, rnorm (n2, mean = mu_2, sd = sqrt (var2))) )
  pvalues_eq <- c()
  pvalues_noteq <- c()
  for (col in 1:ncol(A_sample)) {
  pvalues_eq <-append (pvalues_eq, t.test (A_sample[,col], B_sample[,col], var.equal=TRUE, 
                                           conf.level = 0.95)$p.value )
  pvalues_noteq <-append( pvalues_noteq, t.test (A_sample[,col], B_sample[,col], var.equal=FALSE, 
                                                 conf.level = 0.95)$p.value )}
  pwr_eq_sigma <- append (pwr_eq_sigma, sum (pvalues_eq < alpha)/N)
  pwr_noteq_sigma <-append (pwr_noteq_sigma,  sum (pvalues_noteq < alpha)/N )
  pwr_eq_sigma
  pwr_noteq_sigma
  
  mu_1 = 5
  mu_2 = 0
  mu1_mu2 = 5
  A_sample <- as.data.frame ( replicate (N, rnorm (n1, mean = mu_1, sd = sqrt (var1))) )
  B_sample <- as.data.frame ( replicate (N, rnorm (n2, mean = mu_2, sd = sqrt (var2))) )
  pvalues_eq <- c()
  pvalues_noteq <- c()
  for (col in 1:ncol(A_sample)) {
  pvalues_eq <-append (pvalues_eq, t.test (A_sample[,col], B_sample[,col], var.equal=TRUE, 
                                           conf.level = 0.95)$p.value )
  pvalues_noteq <-append( pvalues_noteq, t.test (A_sample[,col], B_sample[,col], var.equal=FALSE, 
                                                 conf.level = 0.95)$p.value )}
  pwr_eq_sigma <- append (pwr_eq_sigma, sum (pvalues_eq < alpha)/N)
  pwr_noteq_sigma <-append (pwr_noteq_sigma,  sum (pvalues_noteq < alpha)/N )
  pwr_eq_sigma
  pwr_noteq_sigma
  
  return (as.data.frame(cbind (pwr_eq_sigma, pwr_noteq_sigma)))
}

#Added mu1 - mu2 = 0
res <- cbind (c(-5, -1, 0, 1, 5), get_power(10, 10, 1, 1), get_power(10, 10, 3, 1), get_power(10, 10, 9, 1))
names(res)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res
ggplot ()+ geom_line (data = res, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
          geom_line (data = res, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
          geom_line (data = res, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
          geom_line (data = res, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
          geom_line (data = res, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
          geom_line (data = res, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")
  
```

The rest of the combinations. I messed with some of the naming just to keep things seperate for each
```{r}
res1025 <- cbind (c(-5, -1, 0, 1, 5), get_power(10, 25, 1, 1), get_power(10, 25, 3, 1), get_power(10, 25, 9, 1))
names(res1025)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res1025
ggplot ()+ geom_line (data = res1025, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res1025, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res1025, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res1025, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res1025, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res1025, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

res1060 <- cbind (c(-5, -1, 0, 1, 5), get_power(10, 60, 1, 1), get_power(10, 10, 3, 1), get_power(10, 60, 9, 1))
names(res1060)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res1060
ggplot ()+ geom_line (data = res1060, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res1060, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res1060, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res1060, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res1060, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res1060, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

res2510 <- cbind (c(-5, -1, 0, 1, 5), get_power(25, 10, 1, 1), get_power(25, 10, 3, 1), get_power(25, 10, 9, 1))
names(res2510)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res2510
ggplot ()+ geom_line (data = res2510, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res2510, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res2510, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res2510, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res2510, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res2510, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

res6010 <- cbind (c(-5, -1, 0, 1, 5), get_power(60, 10, 1, 1), get_power(60, 10, 3, 1), get_power(60, 10, 9, 1))
names(res6010)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res6010
ggplot ()+ geom_line (data = res6010, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res6010, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res6010, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res6010, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res6010, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res6010, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

res2525 <- cbind (c(-5, -1, 0, 1, 5), get_power(25, 25, 1, 1), get_power(25, 25, 3, 1), get_power(25, 25, 9, 1))
names(res2525)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res2525
ggplot ()+ geom_line (data = res2525, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res2525, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res2525, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res2525, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res2525, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res2525, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

res2560 <- cbind (c(-5, -1, 0, 1, 5), get_power(25, 60, 1, 1), get_power(25, 60, 3, 1), get_power(25, 60, 9, 1))
names(res2560)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res2560
ggplot ()+ geom_line (data = res2560, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res2560, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res2560, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res2560, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res2560, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res2560, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

res6025 <- cbind (c(-5, -1, 0, 1, 5), get_power(60, 25, 1, 1), get_power(60, 25, 3, 1), get_power(60, 25, 9, 1))
names(res6025)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res6025
ggplot ()+ geom_line (data = res6025, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res6025, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res6025, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res6025, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res6025, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res6025, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

res6060 <- cbind (c(-5, -1, 0, 1, 5), get_power(60, 60, 1, 1), get_power(60, 60, 3, 1), get_power(60, 60, 9, 1))
names(res6060)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", "pwr_eq_sigma1_3", "pwr_noteq_sigma1_3", "pwr_eq_sigma1_9", "pwr_noteq_sigma1_9")
res6060
ggplot ()+ geom_line (data = res6060, aes (x = mu_dif, y = pwr_eq_sigma1_1), color = "blue") +
  geom_line (data = res6060, aes (x = mu_dif, y = pwr_noteq_sigma1_1), color = "red") +
  geom_line (data = res6060, aes (x = mu_dif, y = pwr_eq_sigma1_3), color = "green") +
  geom_line (data = res6060, aes (x = mu_dif, y = pwr_noteq_sigma1_3), color = "black")+
  geom_line (data = res6060, aes (x = mu_dif, y = pwr_eq_sigma1_9), color = "grey") +
  geom_line (data = res6060, aes (x = mu_dif, y = pwr_noteq_sigma1_9), color = "brown")

```

```{r}
alpha = 0.05
N = 1000
n1 = c(10,25,60) 
n2 = c(10,25,60)
var1 = c(1,3,9)
var2 = 1
mu1 = 0 
mu2 = c(-5,-1,0,1,5)

for (d in 1:length(n2)){
for (c in 1:length(n1)){

for (b in 1:length(var1)){
  
tmp_pwr <- c()
tmp_mu_diff <- c ()

tmp_pwr2 <- c()
tmp_mu_diff2 <- c()

for (a in 1:length(mu2)){

#Equal Variances
sample1<-replicate(N, rnorm(n1[c], mean=mu1, sd=sqrt(var1[b])))
sample2<-replicate(N, rnorm(n2[d], mean=mu2[a], sd=sqrt(var2)))

pValues<-c()

for (i in 1:N){
  pValues[i]<- t.test(sample1[,i], sample2[,i], var.equal = TRUE, conf.level = 0.95)$p.value
}

accept <- 0
cutoff <- alpha

for (i in 1:N){
  if (pValues[i] <= cutoff){
    accept <- accept + 1
  }
}

pwr <- accept/N

tmp_pwr <- append  (tmp_pwr, pwr)

#Unequal Variances
pValues<-c()

for (i in 1:N){
  pValues[i]<- t.test(sample1[,i], sample2[,i], var.equal = FALSE, conf.level = 0.95)$p.value
}

accept <- 0
cutoff <- alpha

for (i in 1:N){
  if (pValues[i] <= cutoff){
    accept <- accept + 1
  }
}

pwr2 <- accept/N

tmp_pwr2 <- append  (tmp_pwr2, pwr2)

}#end for a

title<-c("equal variance",var1[b], n1[c], n2[d])
Title<-toString(title)
plot(mu2, tmp_pwr, type = "l", main = Title)

title2<-c("unequal variance",var1[b], n1[c], n2[d])
Title2<-toString(title2)
plot(mu2, tmp_pwr2, type = "l", main = Title2)

}#end for b
}#end for c
}#end for d
