---
title: "Group 5, R Project 3"
author: "Kera Whitley, Laura Mathews, Nataliya Peshekhodko"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'

abstract: "This project contains a hypothesis test for the observed data and a simulation study to compare the two hypothesis testing procedures."

output: pdf_document
---
```{r message=FALSE}
library(dplyr)
library (ggplot2)
library (gridExtra)
library (EnvStats)
library(ggpubr)
```

# First part: hypothesis test for the observed data


## 1.1. Read data
Read the data from txt file and save it in dataframe.

```{r}
data <- read.delim("mpg.txt", header=TRUE, sep=" ")
```


## 1.2. Two sample t-test

Conduct two sample t-test twice: with equal variance and with unequal veriance.
T-test with equal variance: 

```{r}
equal.var.test <- t.test( MPG~Country, data = data, var.equal=TRUE, conf.level=0.95)
equal.var.test
```

Based on the equal variance t.test **p-value=`r equal.var.test$p.value`** is small (less then 5% cut off), we can reject the null hypothesis. 

T-test with unequal variance: 

```{r}
unequal.var.test <- t.test( MPG~Country, data = data, var.equal=FALSE, conf.level = 0.95)
unequal.var.test
```

Based on the unequal variance t.test **p-value=`r unequal.var.test$p.value`** is small (less then 5% cut off), we also can reject the null hypothesis. 

## 1.3. How well the normality assumption met by the data

To test data for normality we will use 3 methods:

- Compare histogram of the data to a normal probability curve for the data
- Quantile-quantile plot
- Goodness of fit

First, split up data in two data sets based on the country. 
```{r message=FALSE}
us <- filter(data, Country == "US")
japan <- filter(data, Country == "Japan")
```

Plot histogram with overlaping normal curve and quantile-quantile plot for the US data.

```{r  message=FALSE}
# Plot histogram
hist_us<- ggplot (us, aes(MPG))+geom_histogram(aes(y=..density..),colour="black", fill = "white") + 
  ggtitle("US") + stat_function(fun=dnorm, args = list (mean = mean(us$MPG), sd=sd(us$MPG)), 
                                  colour = "red")
# Plot quantile-quantile plot
qq_us <- ggplot(us, aes(sample = MPG))+stat_qq()+stat_qq_line()
# Arrange two plots in one grid
grid.arrange(hist_us, qq_us, ncol=1, nrow=2)
```

Plot histogram with overlaping normal curve and quantile-quantile plot for the Japan data.

```{r  message=FALSE}
# Plot histogram
hist_jpn <- ggplot (japan, aes(MPG))+geom_histogram(aes(y=..density..),colour="black", fill = "white") + 
  ggtitle ("Japan") + stat_function(fun=dnorm, args = list (mean = mean(japan$MPG), sd=sd(japan$MPG)), 
                                    colour = "red")
# Plot quantile-quantile plot
qq_jpn <- ggplot(japan, aes(sample = MPG))+stat_qq()+stat_qq_line()
# Arrange two plots in one grid
grid.arrange(hist_jpn, qq_jpn, ncol=1, nrow=2)
```

Based on the graphical methods, data does not fit normality well.

Apply goodness of fit test to the data:

```{r}
# Goodness of fit test
gofTest(us, distribution = "norm")
gofTest(japan, distribution = "norm")
```

Small p-values for both data mean that we can reject the null hypothesis (H0 - the data is consistent with normal distribution) and both data sets 
are not consistent with Normal distribution.

## 1.4. Preferable test for selected dataset

Given that the data is violate the normality assumption, there are two options:
1. We can transform our data so that the data becomes normally ditributed
2. Run non-parametric test (Mann-Whitney U test) that does not require the assumption of normality.


# Second part: simulation study

## Purpose

The goal of this simulation - analyze performance of the two tests in terms of controlling $\alpha$ and having larger power 
while varying true variances, sample sizes and true means differences. We will get empirical power by simulation as a function of different 
combinations of $\mu$ differences, different variances and sample size.

## Design

During this simulation next set of combinations will be used:

$\sigma_1^2$ = 1,3,9   $\sigma_2^2$ = 1 \
$n_1$ = 10, 25, 60     $n_2$=10,25,60   \
$\mu_1-\mu_2$ = -5, -1, 0, 1, 5         \
$N$=10000                               \

Under each combination we will generate a sample under $H_A$ N times. Then we will perform t.tests for equal and unequal variances and 
calculate number of times $H_0$ is rejected. Empirical power for both tests will be calculated and analyzed.

Since we need to simulate data as a function of different combination of several parameters and make some manipulations with this data, a function for power calculations is created.

```{r}
get_power <- function (n1, n2) {
  
  alpha = 0.05
  N = 1000
  mu_1 = -5
  mu_2 = 0
  mu1_mu2 = -5
  
  #mu_diffs <- c (-5, -1, 0, 1, 5)
  mu_1_all <- c (-5, -1, 0, 1, 5)
  pwr_eq_sigma <- c()
  pwr_noteq_sigma <- c()
  res <- data.frame(cbind(mu_1_all))
  var_all <- c (1,3,9)
  
  for (k in 1:length (var_all)) {
    
     pwr_eq_sigma <- c()
     pwr_noteq_sigma <- c()
     
     for (j in 1:length(mu_1_all)) {
     pvalues_eq <- c()
     pvalues_noteq <- c()
    
     A_sample <- as.data.frame ( replicate (N, rnorm (n1, mean = mu_1_all[j], 
                                                      sd = sqrt (var_all[k]))) )
     B_sample <- as.data.frame ( replicate (N, rnorm (n2, mean = 0, sd = sqrt (1))) )
        for (col in 1:ncol(A_sample)) {
           pvalues_eq <-append (pvalues_eq, t.test (A_sample[,col], B_sample[,col], 
                                            var.equal=TRUE, conf.level = 0.95)$p.value )
           pvalues_noteq <-append( pvalues_noteq, t.test (A_sample[,col], B_sample[,col], 
                                            var.equal=FALSE, conf.level = 0.95)$p.value ) }
           pwr_eq_sigma <- append (pwr_eq_sigma, sum (pvalues_eq < alpha)/N)
           pwr_noteq_sigma <-append (pwr_noteq_sigma,  sum (pvalues_noteq < alpha)/N )
        }
        res <- cbind(res, as.data.frame(cbind(pwr_eq_sigma,  pwr_noteq_sigma)))
  }
  names(res)<- c ("mu_dif", "pwr_eq_sigma1_1", "pwr_noteq_sigma1_1", 
                  "pwr_eq_sigma3_1", "pwr_noteq_sigma3_1", "pwr_eq_sigma9_1",
                  "pwr_noteq_sigma9_1")
  return (res)
}
```

## Results


```{r}
data_n10_n10 <-get_power(10, 10)
data_n10_n25 <-get_power(10, 25)
data_n10_n60 <-get_power(10, 60)
data_n25_n10 <-get_power(25, 10)
data_n25_n25 <-get_power(25, 25)
data_n25_n60 <-get_power(25, 60)
data_n60_n10 <-get_power(60, 10)
data_n60_n25 <-get_power(60, 25)
data_n60_n60 <-get_power(60, 60)
```



```{r}
plot10_10<-ggplot()+geom_line (data = data_n10_n10, aes (x = mu_dif, 
                                y = pwr_eq_sigma1_1, color = "Var1=1, Var2=1")) +
geom_line (data = data_n10_n10, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                                  color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n10_n10, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                                  color = "Var1=3, Var2=1")) +
geom_line (data = data_n10_n10, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                                  color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n10_n10, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                                  color = "Var1=9, Var2=1")) +
geom_line (data = data_n10_n10, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                                     color = "Var1=9, Var2=1"), linetype="dotted") +
scale_color_discrete("n1 = 10, n2 = 10")+xlab("Mu1-Mu2")+ylab("Power") 
 
plot10_25<- ggplot ()+ geom_line (data = data_n10_n25, aes (x = mu_dif, 
                      y = pwr_eq_sigma1_1, color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n10_n25, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                      color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n10_n25, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                      color = "Var1=3, Var2=1")) +
geom_line (data = data_n10_n25, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                      color = "Var1=3, Var2=1"),  
           linetype="dotted")+
geom_line (data = data_n10_n25, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                      color = "Var1=9, Var2=1")) +
geom_line (data = data_n10_n25, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                      color = "Var1=9, Var2=1"),  
           linetype="dotted") +
scale_color_discrete("n1 = 10, n2 = 25")+xlab("Mu1-Mu2")+ylab("Power") 

plot10_60<- ggplot ()+ geom_line (data = data_n10_n60, aes (x = mu_dif, 
                       y = pwr_eq_sigma1_1,color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n10_n60, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                       color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n10_n60, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                       color = "Var1=3, Var2=1")) +
geom_line (data = data_n10_n60, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                       color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n10_n60, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                       color = "Var1=9, Var2=1")) +
geom_line (data = data_n10_n60, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                       color = "Var1=9, Var2=1"), linetype="dotted") +
scale_color_discrete("n1 = 10, n2 = 60")+xlab("Mu1-Mu2")+ylab("Power") 

plot25_10 <-ggplot ()+ geom_line (data = data_n25_n10, aes (x = mu_dif, y = pwr_eq_sigma1_1,  
                       color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n25_n10, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                       color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n25_n10, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                       color = "Var1=3, Var2=1")) +
geom_line (data = data_n25_n10, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                       color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n25_n10, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                       color = "Var1=9, Var2=1")) +
geom_line (data = data_n25_n10, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                       color = "Var1=9, Var2=1"),  linetype="dotted") +
scale_color_discrete("n1 = 25, n2 = 10")+xlab("Mu1-Mu2")+ylab("Power") 

plot25_25<-ggplot ()+ geom_line (data = data_n25_n25, aes (x = mu_dif, y = pwr_eq_sigma1_1,  
                      color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n25_n25, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                      color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n25_n25, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                      color = "Var1=3, Var2=1")) +
geom_line (data = data_n25_n25, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                      color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n25_n25, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                      color = "Var1=9, Var2=1")) +
geom_line (data = data_n25_n25, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                      color = "Var1=9, Var2=1"),  linetype="dotted") +
scale_color_discrete("n1 = 25, n2 = 25")+xlab("Mu1-Mu2")+ylab("Power") 

plot25_60<-ggplot ()+ geom_line (data = data_n25_n60, aes (x = mu_dif, y = pwr_eq_sigma1_1,  
                      color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n25_n60, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                      color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n25_n60, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                      color = "Var1=3, Var2=1")) +
geom_line (data = data_n25_n60, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                      color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n25_n60, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                      color = "Var1=9, Var2=1")) +
geom_line (data = data_n25_n60, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                      color = "Var1=9, Var2=1"),  linetype="dotted") +
scale_color_discrete("n1 = 25, n2 = 60")+xlab("Mu1-Mu2")+ylab("Power") 


plot60_10<-ggplot ()+ geom_line (data = data_n60_n10, aes (x = mu_dif, y = pwr_eq_sigma1_1,  
                      color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n60_n10, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                      color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n60_n10, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                      color = "Var1=3, Var2=1")) +
geom_line (data = data_n60_n10, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                      color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n60_n10, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                      color = "Var1=9, Var2=1")) +
geom_line (data = data_n60_n10, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                      color = "Var1=9, Var2=1"),  linetype="dotted") +
scale_color_discrete("n1 = 60, n2 = 10")+xlab("Mu1-Mu2")+ylab("Power") 

plot60_25<-ggplot ()+ geom_line (data = data_n60_n25, aes (x = mu_dif, y = pwr_eq_sigma1_1,  
                      color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n60_n25, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                      color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n60_n25, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                      color = "Var1=3, Var2=1")) +
geom_line (data = data_n60_n25, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                      color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n60_n25, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                      color = "Var1=9, Var2=1")) +
geom_line (data = data_n60_n25, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                      color = "Var1=9, Var2=1"),  linetype="dotted") +
scale_color_discrete("n1 = 60, n2 = 25")+xlab("Mu1-Mu2")+ylab("Power") 

plot60_60 <- ggplot ()+ geom_line (data = data_n60_n60, aes (x = mu_dif, y = pwr_eq_sigma1_1, 
                        color = "Var1=1, Var2 = 1")) +
geom_line (data = data_n60_n60, aes (x = mu_dif, y = pwr_noteq_sigma1_1, 
                        color = "Var1=1, Var2=1"), linetype="dotted") +
geom_line (data = data_n60_n60, aes (x = mu_dif, y = pwr_eq_sigma3_1, 
                        color = "Var1=3, Var2=1")) +
geom_line (data = data_n60_n60, aes (x = mu_dif, y = pwr_noteq_sigma3_1, 
                        color = "Var1=3, Var2=1"),  linetype="dotted")+
geom_line (data = data_n60_n60, aes (x = mu_dif, y = pwr_eq_sigma9_1, 
                        color = "Var1=9, Var2=1")) +
geom_line (data = data_n60_n60, aes (x = mu_dif, y = pwr_noteq_sigma9_1, 
                        color = "Var1=9, Var2=1"),  linetype="dotted") +
scale_color_discrete("n1 = 60, n2 = 60")+xlab("Mu1-Mu2")+ylab("Power") 


grid.arrange(plot10_10, plot10_25, plot10_60, ncol=1, nrow=3, 
left ="Dotted lines - Power for the Unequal Varince Test \n Solid lines - 
Power for Equal Variance Test") 
grid.arrange(plot25_10, plot25_25, plot25_60, ncol=1, nrow=3, 
left ="Dotted lines - Power for the Unequal Varince Test \n Solid lines - 
Power for Equal Variance Test")
grid.arrange(plot60_10, plot60_25, plot60_60, ncol=1, nrow=3, 
left ="Dotted lines - Power for the Unequal Varince Test \n Solid lines - 
Power for Equal Variance Test")
```

## Conclusion

The results from the power analysis indicate that the t.tests take into account the absolute value of the difference in means between the two sample populations both when the variances are equal and not equal. This is shown by the fact that each of the graphs is symmetrical around mu1-mu2=0. 
The relationship between the difference in means and the difference in variance is that as the difference in variances between the sample populations increases it takes and larger absolute value of the difference in means to raise the power. This can be seen for both the equal variance t.test and the unequal t.test, though the uneual test does increase the power faster than the equal one.
The impact that sample size has on the t.tests are that as the sample size increases so does the rate in which power is increased. The tests for both the equal and unequal variances give very similar results when the sample size, n, is equal. The reason for this is that when the n's are equal the fraction for the denominator for the t.tests are much closer to each other as in equal variance: $\sqrt{(sp^2/n_1)+(sp^2/n_2)}$ => $\sqrt{\frac{(s_1^2+s_2^2)}{2n}}$ and unequal variance: $\sqrt{(s_1^2/n_1)+(s_2^2/n_2)}$ => $\sqrt{(s_1^2+s_2^2)/2n)}$ which means that the denominator is equal for each of these. So now the variation we see in the plots for these comparisons are from the variability in the data sets generated with these conditions.