---
title: "Group 5, R Project 3"
author: "Kera Whitley, Laura Mathews, Nataliya Peshekhodko"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'

abstract: "This is abstarct."

output: pdf_document
---
```{r message=FALSE}
library(dplyr)
```

# First part


## 1.1. Read data
Read the data from txt file.

```{r}
data <- read.delim("mpg.txt", header=TRUE, sep=" ")
```


## 1.2. Two sample t-test

T-test with equal variance 

```{r}
equal.var.test <- t.test( MPG~Country, data = data, var.equal=TRUE, conf.level=0.95)
equal.var.test
```

Based on the equal variance t.test **p-value=`r equal.var.test$p.value`**, we can reject the null hypothesis. 

T-test with unequal variance: 

```{r}
unequal.var.test <- t.test( MPG~Country, data = data, var.equal=FALSE, conf.level = 0.95)
unequal.var.test
```

Based on the unequal variance t.test **p-value=`r unequal.var.test$p.value`**, we also can reject the null hypothesis. 

## 1.3. How well the normality assumption met by the data

```{r}
hist(filter(data, Country == "US")$MPG)
hist(filter(data, Country == "Japan")$MPG)
```


## 1.4. Preferable test for selected dataset



# Second part

