---
title: "Group 5, R Project 3"
author: "Kera Whitley, Laura Mathews, Nataliya Peshekhodko"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'

output: pdf_document
---
```{r message=FALSE}
library(dplyr)
library (ggplot2)
library (gridExtra)
library (EnvStats)
```

# First part: hypothesis test for the observed data

## 1.1. Read data
Read the data from txt file and save it in dataframe.

```{r}
data <- read.delim("mpg.txt", header=TRUE, sep=" ")
```


## 1.2. Two sample t-test

Conduct two sample t-test twice: with equal variance and with unequal veriance.
T-test with equal variance: 

```{r}
# Apply equal variance t-test
equal.var.test <- t.test( MPG~Country, data = data, var.equal=TRUE, conf.level=0.95)
equal.var.test
```

Based on the equal variance t.test **p-value=`r equal.var.test$p.value`** is small (less then 5% cut off), we can reject the null hypothesis. 

T-test with unequal variance: 

```{r}
# Apply unequal variance t-test
unequal.var.test <- t.test( MPG~Country, data = data, var.equal=FALSE, conf.level = 0.95)
unequal.var.test
```

Based on the unequal variance t.test **p-value=`r unequal.var.test$p.value`** is small (less then 5% cut off), we also can reject the null hypothesis. 

## 1.3. How well the normality assumption met by the data

To test data for normality we will use 3 methods:

- Compare histogram of the data to a normal probability curve for the data
- Quantile-quantile plot
- Goodness of fit

First, split up data in two data sets based on the country. 
```{r message=FALSE}
# Split up data based on the country name
us <- filter(data, Country == "US")
japan <- filter(data, Country == "Japan")
```

Plot histogram with overlaping normal curve and quantile-quantile plot for the US data.

```{r  message=FALSE}
# Plot histogram
hist_us<- ggplot (us, aes(MPG))+geom_histogram(aes(y=..density..),colour="black", fill = "white") + 
  ggtitle("US") + stat_function(fun=dnorm, args = list (mean = mean(us$MPG), sd=sd(us$MPG)), 
                                  colour = "red")
# Plot quantile-quantile plot
qq_us <- ggplot(us, aes(sample = MPG))+stat_qq()+stat_qq_line()
# Arrange two plots in one grid
grid.arrange(hist_us, qq_us, ncol=1, nrow=2)
```

Plot histogram with overlaping normal curve and quantile-quantile plot for the Japan data.

```{r  message=FALSE}
# Plot histogram
hist_jpn <- ggplot (japan, aes(MPG))+geom_histogram(aes(y=..density..),colour="black", fill = "white") + 
  ggtitle ("Japan") + stat_function(fun=dnorm, args = list (mean = mean(japan$MPG), sd=sd(japan$MPG)), 
                                    colour = "red")
# Plot quantile-quantile plot
qq_jpn <- ggplot(japan, aes(sample = MPG))+stat_qq()+stat_qq_line()
# Arrange two plots in one grid
grid.arrange(hist_jpn, qq_jpn, ncol=1, nrow=2)
```

Based on the graphical methods, data does not fit normality well.

Apply goodness of fit test to the data:

```{r}
# Goodness of fit test for us and japan data
gofTest(us, distribution = "norm")
gofTest(japan, distribution = "norm")
```

Small p-values for both data mean that we can reject the null hypothesis (H0 - the data is consistent with normal distribution) and both data sets 
are not consistent with Normal distribution.

## 1.4. Preferable test for selected dataset

Our samples have unequal sizes, so unequal variance test is more preferable here.

We onserved that the data is violate the normality assumption, there are two options:

+ We can transform our data so that the data becomes normally ditributed
+ Run non-parametric test (Mann-Whitney U test) that does not require the assumption of normality.

